---
title: 'BIOS 522: Project 1'
author: "Falcons Group"
urlcolor: blue
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
  word_document: default
---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,          # don't show code
  warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  fig.align = "center"   # center figures
)
library(flexsurv)
library(ggplot2)
library(mice)
set.seed(1)             # make random results reproducible
```

\newpage

# Standard Regression techniques questions

* Dr. Blum is interested in survival times of patients and would like to know the impact of treatment, age, and serum bilirubin as a categorical variable (<1.1, 1.1-3.3, and >3.3) on survival.
  + Use a linear model after excluding all censored observations 
  + Use a linear model after treating censored times as death times
  + Use a logistic regression by defining a new outcomes as dead=1 and otherwise (survived or censored) as 0. 
* For each of above models perform univariate and multivariate analyses(for the three covariates above).
* Interpret the estimates of coefficients of treatment, age, and serum bilirubin regardless of their significance. 
* Now comment on the appropriateness of the data analyses Dr. Blum suggested. A critique. 

# Survival Regression techniques questions

* Now perform a parametric survival analysis (Weibull) and conduct the same univariate and multivariate analyses. Report same results for interpreting coefficients. 
* Dr. Blum wants to know how to derive the i) estimates ii) standard errors of the coefficients that R outputs from the regression. What is the procedure? How can Dr. Blum recreate them herself? Attach this techinical section as an appendix to the report.

# Analysis of the linear regression
## Data input and cleaning

```{r}
# Reading in data
data <- read.table("data.csv", 
                   quote="\"", comment.char="")

# Renaming variables
colnames(data) <- c("caseid", "time", "status", "drug", "age", "sex", "ascites",
                    "hepatomegaly", "spiders", "edema", "bilirubin", "cholesterol",
                    "albumin", "urine_copper", "alk_phosphatase", "sgot", 
                    "triglicerides", "platelets", "prothrombin", "hist_stage")

# Combining censored cases
data$dead <- ifelse(data$status == 0 | data$status == 1, 0, 1)
table(data$dead)

# Creating catergorical factor for bilirubin
data$bili_cat <- ifelse(data$bilirubin < 1.1, "< 1.1", 
                        ifelse(data$bilirubin >= 1.1 & data$bilirubin <= 3.3, "1.1-3.3", ">3.3"))
data$bili_cat <- factor(data$bili_cat, levels = c("< 1.1", "1.1-3.3", ">3.3"))
table( data$bili_cat)

# Creating age in years
data$ageinyear <- data$age/365.25
```

## Excluding censored observations

### Regression on treatment (drug)

```{r}
# univariate analysis for the treatment (drug)
summary(lm(time ~ drug, data = data[data$dead==1,]))
```
### Regression on age

```{r}
# univariate analysis for age
summary(lm(time ~ ageinyear, data = data[data$dead==1,]))
```
### Regression on serum bilirubin

```{r}
# univariate analysis for serum bilirubin
summary(lm(time ~ bili_cat, data = data[data$dead==1,]))
```
### Regression on all three covariates

```{r}
# multivariate analysis for all three covariates
summary(lm(time ~ drug + ageinyear + bili_cat, data = data[data$dead==1,]))
```

## Treating censored times as death times

### Regression on treatment (drug)

```{r}
# univariate analysis for the treatment (drug)
summary(lm(time ~ drug, data = data))
```

### Regression on age

```{r}
# univariate analysis for age
summary(lm(time ~ ageinyear, data = data))
```

### Regression on serum bilirubin

```{r}
# univariate analysis for serum bilirubin
summary(lm(time ~ bili_cat, data = data))
```

### Regression on all three covariates

```{r}
# multivariate analysis for all three covariates
summary(lm(time ~ drug + ageinyear + bili_cat, data = data))
```

# Analysis of the logistic regression

### Regression on treatment (drug)

```{r}
# univariate analysis for the treatment (drug)
summary(glm(dead ~ drug, family = "binomial", data = data))
```

### Regression on age

```{r}
# univariate analysis for age
summary(glm(dead ~ ageinyear, family = "binomial", data = data))
```

### Regression on serum bilirubin

```{r}
# univariate analysis for serum bilirubin
summary(glm(dead ~ bili_cat, family = "binomial", data = data))
```

### Regression on all three covariates

```{r}
# multivariate analysis for all three covariates
summary(glm(dead ~ drug + ageinyear + bili_cat, family = "binomial", data = data))
```

# Critique of Dr. Blum's analyses

# Parametric survival analysis

## Parametric survival analysis on drug

```{r}
# here we used library "flexsurv" for the analysis
# univariate analysis for the treatment (drug)
flexsurvreg(Surv(time, dead) ~ drug, data = data, dist = "Weibull")
```

## Parametric survival analysis on age

```{r}
# univariate analysis for age
flexsurvreg(Surv(time, dead) ~ ageinyear, data = data, dist = "Weibull")
```

## Parametric survival analysis on serum bilirubin

```{r}
# univariate analysis for serum bilirubin
flexsurvreg(Surv(time, dead) ~ bili_cat, data = data, dist = "Weibull")
```

## Parametric survival analysis on all three covariates

```{r}
# multivariate analysis for all three covariates
flexsurvreg(Surv(time, dead) ~ drug + ageinyear + bili_cat, data = data, dist = "Weibull")
```

## Checking Weibull distribution is a good fit with log-log plot

```{r}
# Create cumulative hazard
data$ch <- nelsonaalen(data, time, dead)
data$drug <- ifelse(data$drug == 1, "treatment", "placebo")
ggplot(data, aes(log(time), log(ch))) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  facet_grid(drug ~ bili_cat) +
  theme_bw() + 
  ylab("Log Cumulative Hazard") + 
  xlab("Log time (in days)") + 
  ggtitle("Check fit data to Weibull Distribution")
```
The plot of log time against the log cumulative hazard is mostly linear in each of the treatment / serum bilirubin groups and deviates mostly at the earlier time points. Due to the linearity within each of the groups, we conclude the Weibull distribution is a good fit for this dataset. It is harder to visualize along with age (our third covariate), but based on this diagnostic it seems Weibull is a good fit regardless.

# Technical Appendix: How to derive coefficient estimates and standard errors from parametric survival analyses

\newpage

# Code

```{r ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
# this R markdown chunk generates a code appendix
```
