---
title: 'BIOS 522: Project 1'
author: "Falcons Group"
urlcolor: blue
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
  word_document: default
---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,          # don't show code
  warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  fig.align = "center"   # center figures
)
library(data.table)
library(ggplot2)

set.seed(1)             # make random results reproducible
```

```{r setup, include=FALSE}
data <- fread("data.csv")

cols <- c("id", "time", "status", "drug", "age", "sex", 
          "is_ascites", "is_hept", "is_spiders", "is_edema", 
          "serum_bilirubin", "serum_chol", "albumin", "urine_copper", 
          "alk_phosphatase", "sgot", "triglicerides", "platelets", 
          "prothrombin", "histologic_stage")

names(data) <- cols

data <- data[, .(id, time, status, drug, age, serum_bilirubin)]
data[, serum_cat := cut(serum_bilirubin, c(0, 1.1, 3.3, max(serum_bilirubin)))]
```

\newpage

# Standard Regression techniques questions

* Dr. Blum is interested in survival times of patients and would like to know the impact of treatment, age, and serum bilirubin as a categorical variable (<1.1, 1.1-3.3, and >3.3) on survival.
  + Use a linear model after excluding all censored observations 
  + Use a linear model after treating censored times as death times
  + Use a logistic regression by defining a new outcomes as dead=1 and otherwise (survived or censored) as 0. 
* For each of above models perform univariate and multivariate analyses(for the three covariates above).
* Interpret the estimates of coefficients of treatment, age, and serum bilirubin regardless of their significance. 
* Now comment on the appropriateness of the data analyses Dr. Blum suggested. A critique. 

# Survival Regression techniques questions

* Now perform a parametric survival analysis (Weibull) and conduct the same univariate and multivariate analyses. Report same results for interpreting coefficients. 
* Dr. Blum wants to know how to derive the i) estimates ii) standard errors of the coefficients that R outputs from the regression. What is the procedure? How can Dr. Blum recreate them herself? Attach this techinical section as an appendix to the report.

# Analysis of the linear regression

## Excluding censored observations

## Treating censored times as death times

# Analysis of the logistic regression

# Critique of Dr. Blum's analyses

# Parametric survival analysis

# Technical Appendix: How to derive coefficient estimates and standard errors from parametric survival analyses

\newpage

# Code

```{r ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
# this R markdown chunk generates a code appendix
```
